---
title: "intRo 3"
subtitle: "R Markdown"
author: "Billy Fryer"
output: pdf_document
---

# R Markdown

R Markdown is probably my favorite thing about R Studio. R Markdown files allow you to save output from code as a pdf, html, a doc and many more file types. For this tutorial, we will mostly be focusing on the pdf output, since that is what we use the most in SAC.

# Getting Started

You must have LaTeX in order to make pdf files using R Markdown. For more information, go to https://bookdown.org/yihui/rmarkdown-cookbook/install-latex.html

To open a new R Markdown File, click the drop down arrow next to the sheet of paper with the green plus on it in the top left. Then select "R Markdown". To actually make the file after typing the code and plain text, hit the "Knit" Button. This will run all of your code and create the file.


# YAML Heading

The YAML (pronounced yam-el) Heading is super important in R Markdown. It is where many of your global options go and is automatically included when opening a new R Markdown Document. The YAML Heading should always be at the beginning of the R Markdown document. The YAML Heading for this document is pasted below:

---
title: "intRo 3"
author: "Billy Fryer"
output: pdf_document
---

As you can see, title and author are self explanatory. You can also include the date using a similar format. For Sports Analytics Club, dating your work is very important when working with one of NCSU's varsity athletic programs so that they know what data is being used. The last one we will discuss right now is the output. There are several different formats that the output can be. Normally I choose to make it a pdf, but there may be times where other formats are more useful. 

# Set Up Chunk

The Set Up Chunk normally goes to the very top of the document, but for this example, it has been moved

```{r setup}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, warning = FALSE, message = FALSE)
```

There are several options that one can select in the setup chunk. There functions are explained below:

echo = if TRUE, it makes the code appear in the output document. Normally, we want this to be FALSE when showing to a team because they don't care about pretty code, they just want results.

include = if TRUE, the code chunk is included in the output. Typically, this is marked as FALSE if echo = TRUE or vice versa.

warning = cuts off warnings if FALSE or prints them when it is TRUE. Typically this is FALSE because we don't want warnings to show up in our output. However, if we are debugging, turning this on can be useful. If your code will not work, the knitting of the document will not happen, so don't worry about that.

message = is pretty much the warning.

# Formatting

Formatting in R Markdown can take place in a variety of ways from headers to spacing to writing equations (which is useful in stats classes, not as much in Sports Analytics Club)

## Headings

Headings occur by putting # in front of the text. For whatever level you want, put that many #s and then a space

# This is a First Level Header

Written "# This is a First Level Header" (without the quotes)

## This is a Second Level Header

Written "## This is a Second Level Header" (without the quotes)


### This is Third Level Header

Written "### This is a Third Level Header" (without the quotes)


## Spacing

Spacing in R Markdown is weird. To space between words is normal, just hit the space bar. However, if you want to space between lines it gets more complicated.

This is an example of using the two spaces.  
Now we are on a new line.

I prefer using a blank line in between lines.

It makes your code longer, but it makes it a lot easier to see.

To do a new page, use the \\newpage command. It typically turns a different color. (It was used to seperate to the next page here)

\newpage

## The \ Character and Writing Equations

The \\ character is a tricky one to deal with. Typically it is used for writing LaTeX commands in R. If for some reason you need to write it (such as writing this tutorial), you would type the character twice for it to print once.

Writing equations in R is a very convenient way to do homework in your statistics classes. You start and end with a \$ and then use LaTeX commands in between. If you don't know how to write in LaTeX, don't worry. Google is very helpful if you search "how to write (character you want) in latex". Sometimes, this also might be the most convenient way to write special characters in R Markdown.An example is below (though you can write equations in the middle of the sentence).

$E(Y) = \beta_0 + \beta_1X_1 + \beta_2X_2 + ...+\beta_nX_n$

The above equation was written: \$E(Y) = \\beta_0 + \\beta_1X_1 + \\beta_2X_2 + ...+\\beta_nX_n\$

## Text Formatting

You can also Bold or Underline text in R Markdown. To **bold** text, surround the text with ** (Two Asterisks), without a space before the first word of the phrase and without a space after the last word of the phrase. To *italicize* text, do the same thing as bolding, but only use * (One Asterisk). There is currently no good way to underline text, so use bolding and italicizing when appropriate.

# Writing Code in R Markdown

This is what makes R Markdown the best: including R Code and Output directly in with Plain text. R Code goes in Code Chunks, which are signified with 3 tick marks \`\`\` (the key above tab and below the esc key on mac and pc). Then you use an open curly bracket { then the name of the language you are using (in our case, r though you can use python or other languages in R Markdown). From there you need to name the code chunk. Typically this name will be descriptive in order to make it easier to find when you let the code sit for a while and then come back to it. Names are technically optional, but named chunks are preferred. When naming your chunks, each chunk needs a distinct name for R Markdown to knit properly (regardless of output type). From there you close the curly bracket } and then type your code. Coding in the R chunk is the same as coding in a R script. When you are finished with the code, do three more tick marks \`\`\` on a separate line.

That was a lot of text, let's write out an example.

\`\`\`{r sample}
library(tidyverse) # Pretty Much Always Need Tidyverse!
library(Lahman) # Bring in Lahman Datasets

lahmanP <- People %>% filter(birthState == "NC") # Only NC Pitchers

lahmanP <- lahmanP %>% select(nameGiven, debut, playerID) # Useful Variables

print(head(lahmanP))
\`\`\`

\newpage

**Let's actually do this!**

```{r real sample}
library(tidyverse) # Pretty Much Always Need Tidyverse!
library(Lahman) # Bring in Lahman Datasets

lahmanP <- People %>% filter(birthState == "NC") # Only NC Pitchers

lahmanP <- lahmanP %>% select(nameGiven, debut) # Useful Variables

print(head(lahmanP)) # Print the first few rows
```

# Conclusion

There are a lot more things to learn about R Markdown; these are only the basics! Hope you found it helpful!